---
title: "2018_midterm_Stroud"
author: "Hannah Stroud"
date: "October 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
###Load the libraries###
library(readr)
library(dplyr)
library(MASS)
library(tidyr)
library(ggplot2)
library(profileModel)
library(brms)
library(tidybayes)
library(bayesplot)
library(rstanarm)
library(tidyverse)
library(reprex)
library(datapasta)
library(dplyr)
```
## 1. Sampling your system (10 points)
####Each of you has a study system your work in and a question of interest. Give an example of one variable that you would sample in order to get a sense of its variation in nature. Describe, in detail, how you would sample for the population of that variable in order to understand its distribution. Questions to consider include, but are not limited to: Just what is your sample versus your population? What would your sampling design be? Why would you design it that particular way? What are potential confounding influences of both sampling technique and sample design that you need to be careful to avoid? What statistical distribution might the variable take, and why?

I don't deal with natural system and I'm not sure how to apply this question to what I study. See extra credit in attempt to make up missed points.  

## 2. Let's get philosophical. (10 points)
####Are you a frequentist, likelihoodist, or Bayesian? Why? Include in your answer why you prefer the inferential tools (e.g. confidence intervals, test statistics, posterior probabilities, etc.) of your chosen worldview and why you do not like the ones of the other one. This includes defining just what those different tools mean! extra credit for citing and discussing outside sources - one point per source/point

I am a Bayesian. I think there is a communication issue with frequentist statistics to those who don't deal with them and I feel like while they can be used for good, the p-value has been used without much thought at times, only relying on the 0.05 rule. I have found the idea "We teach it because it’s what we do; we do it because it’s what we teach" to be too often true(Cobb qtd. in Wasserstein & Lazar 2016). I like that Bayesian inference "provides an explicit expression of the amount of uncertainty" in estimates, which is a lot more useful than saying you rejected the null or not (Ellison 1996). It's also  As Ellison also points out, frequentist assumptions rely on true random sampling and a true fixed value for each parameter of interest- and neither of those seem to fit well for me (1996). Another big statistics and science communication element with Frequentist statistics relates to confidence intervals. Confidence intervals in frequentists statistics are frequently misinterpreted, because people read them as 95% confident that the true mean lies within the interval but really they mean "Under repeated sampling, the true mean will be inside the confidence interval 95% of the time" ("Abraham Lincoln and confidence intervals" https://andrewgelman.com/2016/11/23/abraham-lincoln-confidence-intervals/).  I do find times where p-values, t-tests and other frequentist statistics are useful, but I feel like overall they aren't used in the most useful way as the ASA guidelines from 2016 recommend. 

Likelihood is more intuitive to mean than using t tests and p-values, but is still grounded in frequentist thought. Likelihood value only "gains meaning when compared the likelihoods of other possible values" (Whitlock & Schluter , p. 656). It samples all possible parameters within a range to find the one most likely to give the data you found. It is more versatile than least squares but falls short with complicated models. 

Intuitively, the concept behind Bayesian statistics just fits better with how I view the world because it presents the confidence in a hypothesis given the data when frequentist and likelihood give the probability of the data given a hypothesis. I really like the idea and simplicity of priors and posteriors. Bayesian is a beautiful statistical circle of life: "we acquire data..update the posterior distribution...This updated distribution then serves as the prior" for future models.  (Seeing Theory: https://seeing-theory.brown.edu/bayesian-inference/index.html#section1). Priors are really useful to me- they allow us to utilize the knowledge we already have about what we are studying and improve upon it, and incorporating priors just seems like a better idea than relying on random sampling that is never truly perfectly random or normally distributed. Bayesian modeling is also more versatile. We can model using Markov Chain Monte Carlo analysis, which allows us to model non-gaussian distributions and hierarchal relationships. Overall, while the code is longer, the Bayesian tools seem better applied to wide range of models and easier to communicate results.  
## 3. Power (20 points)
#### We have a lot of aspects of the sample of data that we collect which can alter the power of our linear regressions.

####Slope
####Intercept
####Residual variance
####Sample Size
####Range of X values
#### Choose three of the above properties and demonstrate how they alter power of an F-test from a linear regression using at least three different alpha levels (more if you want!) As a baseline of the parameters, let’s use the information from the seal data:

slope = 0.00237, intercept=115.767, sigma = 5.6805, range of seal ages = 958 to 8353, or, if you prefer, seal ages N(3730.246, 1293.485). Your call what distribution to use for seal age simulation.

Extra Credit 2 Choose just one of the above elements to vary. Using likelihood to fit models, repeat your power analysis for a chi-square likelihood ratio test. You can use glm(), bbmle or some other means of fitting and obtaining a LRT at your discretion. 5 points.  

```{r}

sim_parameters <- data.frame(intercept = c(100, 115,120,125),
                      slope= 0.00237) %>%
  crossing(residual_sd = 3:8) %>%
  crossing(sample_size = 5:15) %>%    #looking at sample size, intercept, residual
  #set up sampling 
  group_by(intercept, slope, residual_sd, sample_size) %>%
  expand(reps= 1:sample_size) %>%
  ungroup() %>%
  #replicate for simulations 
  crossing(sim= 1:100) %>% 
  mutate(age.days= runif(n(), 958, 8353)) %>% 
           mutate(length.cm = rnorm(n(), intercept + slope* age.days,
                                    residual_sd)) %>%
  ungroup()
  ##fit models##   
  
 lim_fit <- function(sim_parameters){ 
  fit <- sim_parameters %>% 
    group_by(sim, intercept, slope, residual_sd, sample_size) %>%
    nest() %>%
    mutate(mod = purrr:::map(data, ~lm(length.cm ~ age.days, data=.))) %>%
    mutate(coefs = purrr::map(mod, ~broom::tidy(.))) %>%
    unnest(coefs) %>%
    ungroup() %>%
    filter(term == "age.days")
  }
sim_fit <- lim_fit(sim_parameters= sim_parameters)  

#power for intercept change 
 pow_intercept <- sim_fit %>% 
   crossing(alpha= c(0.001, 0.01, 0.05, 0.1))%>%
   group_by(intercept, alpha)%>% 
   mutate(power=1-sum(p.value>alpha)/n()) %>%
   ungroup()

###Rinse and repeat for residual sd, and sample size
pow_resid <- sim_fit %>% 
  crossing(alpha=c(0.001, 0.01, 0.05, 0.1)) %>%
  group_by(residual_sd, alpha) %>%
  mutate(power= 1- sum(p.value>alpha)/n()) %>%
  ungroup()

pow_samp_siz <- sim_fit %>% 
  crossing(alpha= c(0.001, 0.01, 0.05, 0.1))%>%
  group_by(sample_size, alpha) %>% 
  mutate(power=  1- sum(p.value>alpha)/n()) %>%
  ungroup()

#plot 
inter_plot <- ggplot(data = pow_intercept, 
       mapping = aes(x = intercept, y = power, color = factor(alpha))) +
  geom_point() +
  geom_line() +
  theme_bw()
resid_plot<- ggplot(data = pow_resid, 
       mapping = aes(x = residual_sd, y = power, color = factor(alpha))) +
  geom_point() +
  geom_line() +
  theme_bw()
samp_siz_plot <- ggplot(data = pow_samp_siz, 
       mapping = aes(x = sample_size, y = power, color = factor(alpha))) +
  geom_point() +
  geom_line() +
  theme_bw()

#visualize#
samp_siz_plot
resid_plot
inter_plot

```
Change in intercept did not have a significant effect on power, increasing residual variance had a negative effect on power and larger sample sizes increased power. 

```{r extra credit}
#EC: Extra credit 1 - test whether the distribution of ages alters power: 3 points

sim_log<- data.frame(intercept = 115.767,
                      slope= 0.00237, residual_sd = 5.6805) %>%
 crossing(sample_size = 5:10) %>%  
group_by(intercept, slope, residual_sd, sample_size) %>%
    expand(reps= 1:sample_size) %>%
    ungroup() %>%
    #replicate for simulations 
    crossing(sim= 1:100) %>% 
    mutate(age.days= rlnorm(n(), log(3730.246), log(1293.485))) %>% 
    mutate(length.cm = rnorm(n(), intercept + slope* age.days,
                             residual_sd)) %>%
    ungroup()
 
#fit
log_fit <- sim_log %>% 
    group_by(sim, sample_size) %>%
    nest() %>%
    mutate(mod = purrr:::map(data, ~lm(length.cm ~ age.days, data=.))) %>%
    mutate(coefs = purrr::map(mod, ~broom::tidy(.))) %>%
    unnest(coefs) %>%
    ungroup() %>%
    filter(term == "age.days")
#power 
pow_age_distrib <- log_fit %>% 
  crossing(alpha= c(0.001, 0.01, 0.05, 0.1))%>%
  group_by(sample_size, alpha) %>% 
  mutate(power=  1- sum(p.value>alpha)/n()) %>%
  ungroup()
log_dis_plot <- ggplot(data = pow_age_distrib, 
       mapping = aes(x = sample_size, y = power, color = factor(alpha))) +
  geom_point() +
  geom_line() +
  theme_bw()
log_dis_plot
samp_siz_plot
```
Power is higher with a log distribution.    
## 4. I’ve referenced the following figure a few times. I’d like you to demonstrate your understanding of Bayes Theorem by hand showing what the probability of the sun exploding is given the data. Assume that your prior probability that the sun explodes is p(Sun Explodes) = 0.0001. The rest of the information you need is in the cartoon!

```{r Bayesian Theory}
P_E <- 0.0001  
P_N <- 1- P_E #probabilit it doesn't explode
P_Y_E <- 35/36    #prob it says the sun explodes given that it did explode
P_Y_NE <- 1/36  #prob it says yes given the sun doesn explode (ie the prob of it lying)

#Prob of explostion give true= P(Y|E)* (P(E))/ P(Y)
#Prob of it being true P(Y)= P(Y|E) P(E) + P(Y|N) * P(N)


#our equation
#P(Exp |Yes) <- P(yes|explodes)*p(Explodes)/Prob(Yes)
#Prob(Y)= P(Says Ex | explodes)*P(E) + P(Says ex|doesnt)* P(N)

P_Y <- P_Y_E*P_E + P_Y_NE*P_N

P_exp_given_data <- (P_Y_E* P_E)/P_Y
P_exp_given_data
```
We want to know the probability of the sun exploding given the machine says yes P(E| Y)
The probability the sun explodes given the data is 0.0035.  

##5 Quailing at the Prospect of Linear Models
I’d like us to walk through the three different ‘engines’ that we have learned about to fit linear models. To motivate this, we’ll look at Burness et al.’s 2012 study "Post-hatch heat warms adult beaks: irreversible physiological plasticity in Japanese quail 

```{r cache=TRUE, warning=FALSE, message=FALSE}
library(readr)
library(profileModel)
morph <- read.csv("./Data/Morphology data.csv")
morph <- morph %>% na.omit()    #getting rid of NAs
#visualize data 
ggplot(data= morph, aes(x=Tarsus..mm., y=Culmen..mm.))+ geom_point()+
  stat_smooth(method= "lm")
```

###5.1 Three fits (10 points)
To begin with, I’d like you to fit the relationship that describes how Tarsus (leg) length predicts upper beak (Culmen) length. Fit this relationship using least squares, likelihood, and Bayesian techniques. For each fit, demonstrate that the necessary assumptions have been met. Note, functions used to fit with likelihood and Bayes may or may not behave well when fed NAs. So look out for those errors.
```{r}
#fit with least squares
morph_mod <- lm(morph$Culmen..mm.~ morph$Tarsus..mm.)
#check assumptions 
plot(morph_mod, which=1)   
plot(morph_mod, which=2) #qq-plot looks a little weird on ends 
#tests
anova(morph_mod) #t-test
summary(morph_mod) #High F value, looks good 
```

```{r Likelihood}
#fit with likelihood
morph_glm <- glm(Culmen..mm.~Tarsus..mm., 
                 data=morph,
                 family= gaussian(link="identity"))
#check assumptions 
#assumptions
m_fit <- predict(morph_glm)
m_res <- residuals(morph_glm)

plot(morph_glm, which=1)
plot(morph_glm, which=2)

qqnorm(m_res)
qqline(m_res)   #goodenough
plot(profile(morph_glm, objective= "ordinaryDeviance")) 

#f-tests of model
mor_null <- glm(Culmen..mm. ~ 1, 
                     family = gaussian(link = "identity"),
                     data=morph)

anova(mor_null,morph_glm, test = "LRT") #low p, reject the null
#t-tests of parameters
summary(morph_glm)

```

```{r}
#Bayesian 
morph_bay_mod <- brm(Culmen..mm.~Tarsus..mm.,
                      family = gaussian(link = "identity"),
                      data= morph,
                    file= "./morph_bay_mod")
#test assumptions 
morph_posterior <- posterior_samples(morph_bay_mod, 
                                    add_chain = T)
rhat(morph_bay_mod)#yes! close to 1
mcmc_acf(morph_posterior) #also looks good, convergering  
morph_fit <- fitted(morph_bay_mod) %>% as.data.frame() 
morph_res <- residuals(morph_bay_mod) %>% as.data.frame()
mcmc_trace(morph_posterior)

qqnorm(morph_res$Estimate) #looks good 
pp_check(morph_bay_mod, type= "scatter") #no pathological diversions
pp_check(morph_bay_mod, type = "dens")
pp_check(morph_bay_mod, type = "stat", stat = "mean")
pp_check(morph_bay_mod, type = "stat_2d")
```

5.2 Three interpretations (10 points)
OK, now that we have fits, take a look! Do the coefficients and their associated measures of error in their estimation match? How would we interpret the results from these different analyses differently? Or would we? Note, confint works on lm objects as well.
```{r}
library(tidyverse)
#least squares linear regression
summary(morph_mod)%>% broom::tidy()
confint(morph_mod)

#likelihood   #what about residual se?
morph_glm
confint(morph_glm)

#Baysian 
summary(morph_bay_mod, digits= 4)  #so far so good
```

All three generate a relationship where tarsus predicts culmen length. But they take very different approaches. In least squares, we use test statistics from Frequentist theory where we could confidently reject the null (the probably there was no relationship). 
Likelihood approaches this questsion by looking at the probabiliy of getting our data for all parameters (in a given range). This approach evauluates multiple hypothesis not just comparing ours to a null like above. Like least squares it still looks at the probability fo the data given the hypothesis.  
With a Bayesian approach we can flip the relationship and examine the probability of our hypothesis given the data. It incorporates prior knowledge of the relationship into the model and allows us to assert a degree of belief in the probability of a hypothesis being true.  

###5.3 Everyday I’m Profilin’ (10 points)For your likelihood fit, are your profiles well behaved? For just the slope, use grid sampling to create a profile. You’ll need to write functions for this, and use the results from your glm() fit to provide the reasonable bounds of what you should be profiling over (3SE should do). Is it well behaved? Plot the profile and give the 80% and 95% CI. Verify your results with profileModel.
```{r}
#grid sampling 
lik_fun <- function(slope, intercept, resid_sd){
  #data generating process
 cul_fit <- intercept + slope * morph$Tarsus..mm.
  
  #likelihood
  sum(dnorm(morph$Culmen..mm., cul_fit, resid_sd, log=TRUE))
}

gs_loglik <- crossing(intercept = seq(-0.75, 0.55, .05),
                      slope = seq(0.28,0.48,.01),
                       resid_sd = seq(.5, 1.5, .01)) %>%
  rowwise() %>%
  mutate(logLik = lik_fun(slope, intercept, resid_sd)) %>%
  ungroup()
#filter for max loglik
slope_prof <-gs_loglik %>% group_by(slope) %>%
  filter(logLik == max(logLik)) 

prof_plot <- ggplot(data = slope_prof, aes(x= slope, y=logLik))  + geom_line() +  xlim(0.35,0.39)+ ylim(-1260,-1248) #zoomed in after viewing 
prof_plot  

#95 CI
gs_loglik %>%
filter(logLik > max(logLik) - 1.96) %>% 
  arrange(slope) %>%
  filter(row_number()==1 | row_number()==n())

#80 CI
gs_loglik %>% filter(logLik > max(logLik) - 1.642/2) %>% arrange(slope) %>%
  filter(row_number()==1 | row_number()==n())

prof_plot + geom_hline(yintercept = -1251.6, color="red")+ geom_hline(yintercept= -1250.4, color= "purple")


#compare to profileModel/ sanity check
prof_glm <- profileModel(morph_glm, 
                         objective="ordinaryDeviance",
                         quantile= qchisq(0.80,0.95))
plot(prof_glm)  
```


###5.4 The Power of the Prior (10 points)
####This data set is pretty big. After excluding NAs in the variables we’re interested in, it’s over 766 lines of data! Now, a lot of data can overhwelm a strong prior. But only to a point. Show first that there is enough data here that a prior for the slope with an estimate of 0.4 and a sd of 0.01 is overwhelmed by the data by demonstrating that it produces similar results to our already fit flat prior.
```{r  warning=FALSE, message=FALSE}
# set prior- rstanarm, signif faster 
pr_m <- stan_glm(Culmen..mm. ~ Tarsus..mm.,
                                data = morph,
                                family=gaussian(),
                                prior = normal(0.4,0.01))
#test assumptions 
prm_post <- posterior_samples(pr_m, add_chain= T)
mcmc_acf(prm_post)
rhat(pr_m)  #close to 1
pp_check(pr_m) #converge quickly

summary(pr_m, digits= 4) #0.4 is not in our CI
#intercept is off but slope and residual sd are similar. 
```

####Second, see if a very small sample size (n = 10) would at least include 0.4 in it’s 95% Credible Interval. 
```{r small sample size}
#set  up small sample size df

sm_data <- data.frame(Tarsus=morph$Tarsus..mm.[1:10],
                      Culmen= morph$Culmen..mm. [1:10])


sm_morph_pr <- stan_glm(Culmen~ Tarsus,
                      data = sm_data,
                          family=gaussian(),
                          prior = normal(0.4,0.01))
summary(sm_morph_pr, digits= 4)
posterior_interval(sm_morph_pr)
```
The strong prior influences the small sample size.   

####Last, demonstrate at what sample size that 95% CL first begins to include 0.4 when we have a strong prior. How much data do we really need to overcome our prior belief? Note, it takes a long time to fit these models, so, try a strategy of spacing out the 3-4 sample sizes, and then zoom in on an interesting region.
```{r}
# vary the sample sizes 
g1_data <- data.frame(Tarsus=morph$Tarsus..mm.[1:50],
                      Culmen= morph$Culmen..mm. [1:50])


g1_morph_pr <- stan_glm(Culmen~ Tarsus,
                      data = g1_data,
                          family=gaussian(),
                          prior = normal(0.4,0.01))
posterior_interval(g1_morph_pr)  


g2_data <- data.frame(Tarsus=morph$Tarsus..mm.[1:250],
                      Culmen= morph$Culmen..mm. [1:250])
g2_morph_pr <- stan_glm(Culmen~ Tarsus,
                      data = g2_data,
                          family=gaussian(),
                          prior = normal(0.4,0.01))
posterior_interval(g2_morph_pr)  


g3_data <- data.frame(Tarsus=morph$Tarsus..mm.[1:150],
                      Culmen= morph$Culmen..mm. [1:150])
g3_morph_pr <- stan_glm(Culmen~ Tarsus,
                      data = g3_data,
                          family=gaussian(),
                          prior = normal(0.4,0.01))
posterior_interval(g3_morph_pr)  

g4_data <- data.frame(Tarsus=morph$Tarsus..mm.[1:200],
                      Culmen= morph$Culmen..mm. [1:200])
g4_morph_pr <- stan_glm(Culmen~ Tarsus,
                      data = g4_data,
                          family=gaussian(),
                          prior = normal(0.4,0.01))
posterior_interval(g4_morph_pr) 

g5_data <- data.frame(Tarsus=morph$Tarsus..mm.[1:225],
                      Culmen= morph$Culmen..mm. [1:225])
g5_morph_pr <- stan_glm(Culmen~ Tarsus,
                      data = g5_data,
                          family=gaussian(),
                          prior = normal(0.4,0.01))
posterior_interval(g5_morph_pr)  
```
Somewhere around 225-250 data points the strong prior is overridden.  

###6. Extra CreditMake an election forecast as discussed at https://biol607.github.io/extra.html - but this isn’t just a winner prediction. 1 point for the correct winner. 5 points for correctly predicting the popular vote and being within 10% (3% just for trying!). 5 points for predicting the electoral college and geting no more than 5 states wrong (3 points just for trying). 5 points for predicting the senate races getting no more than 5 states wrong (3 points just for trying). 1 extra point for each vote percentage within your 80% Confidence/Credible Interval. Ditto for the house races.

```{r, MA Senate Seat}
ma_senate <- read_csv("./data/ma_senate_sum_polls.csv")
Warren_wgt <- weighted.mean(x= ma_senate$Warren, y=ma_senate$Size)
Diehl_wg <-weighted.mean(x=ma_senate$Diehl, y=ma_senate$Size)

#congressional re-election rate over all: 90%
sim <-matrix(rbinom(1,500, 0.55 ), nrow=100)


```
Based on polling data, I predict Warren will win with a populat vote of 55.5 percent.  

```{r, Florida Governors Race}
#attempts at a prior based on partisianship
#read in polling data (same polls as used by 538 model)
Fl_gov <- read_csv("./data/florida_gov.csv")
Gillum <- weighted.mean(x=Fl_gov$Gillum, y=Fl_gov$Polled)
DeSantis <- weighted.mean(x=Fl_gov$DeSantis, y=Fl_gov$Polled)
#538 General rule of thumb is partisanship (as defined by presidential elections) voting is 1/3 as strong in governors races as federal 
#FL 2016 
Fl_pres_16 <- 0.49 #voted for republican 


Fl_reg<- 13200872
Fl_rep <- 4661230/Fl_reg  #GOP  
Fl_dem <- 4918415/Fl_reg   #Dems
Fl_min <- 99322/Fl_reg    #minority parties 
Fl_ind <- 3521905/Fl_reg  #no affliation

Fl_partisianship_gov <- 0.33333* 0.49
sum(Fl_gov$Polled)


```
