---
title: "2018_midterm_Stroud"
author: "Hannah Stroud"
date: "October 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Sampling your system (10 points)
####Each of you has a study system your work in and a question of interest. Give an example of one variable that you would sample in order to get a sense of its variation in nature. Describe, in detail, how you would sample for the population of that variable in order to understand its distribution. Questions to consider include, but are not limited to: Just what is your sample versus your population? What would your sampling design be? Why would you design it that particular way? What are potential confounding influences of both sampling technique and sample design that you need to be careful to avoid? What statistical distribution might the variable take, and why?

I don't deal with natural system and I'm not sure how to apply this question to what I study. See extra credit in attempt to make up missed points.  

## 2. Let’s get philosophical. (10 points)
####Are you a frequentist, likelihoodist, or Bayesian? Why? Include in your answer why you prefer the inferential tools (e.g. confidence intervals, test statistics, posterior probabilities, etc.) of your chosen worldview and why you do not like the ones of the other one. This includes defining just what those different tools mean! extra credit for citing and discussing outside sources - one point per source/point

I am a Bayesia. I think there is a communication issue with frequentist statistics to those who don't deal with them and I feel like while they can be used for good, the p-value has been used without much thought at times, only relying on the 0.05 rule. I have found the idea "We teach it because it’s what we do; we do it because it’s what we teach" to be too often true(Cobb qtd. in Wasserstein & Lazar 2016). I like that Bayesian inference "provides an explicit expression of the amount of uncertainty" in estimates, which is a lot more useful than saying you rejected the null or not (Ellison 1996). It's also  As Ellison also points out, frequentist assumptions rely on true random sampling and a true fixed value for each parameter of interest- and neither of those seem to fit well for me (1996). I do find times where p-values, t-tests and other frequentist statistics are useful, but I feel like overall they aren't used in the most useful way as the ASA guidelines from 2016 recommend. 

I'm a Bayesian because I really like the idea and simplicity of priors and posteriors. The core of Bayesian, is a beautiful statistical circle of life: "we aqquire data..update the posterior distribution...This updated distribution then serves as the prior" for future models.  (Seeing Theory: https://seeing-theory.brown.edu/bayesian-inference/index.html#section1). Priors are really useful to me- they allow us to utilize the knowledge we already have about what we are studying and improve upon it. 
## 3. Power (20 points)
#### We have a lot of aspects of the sample of data that we collect which can alter the power of our linear regressions.

####Slope
####Intercept
####Residual variance
####Sample Size
####Range of X values
#### Choose three of the above properties and demonstrate how they alter power of an F-test from a linear regression using at least three different alpha levels (more if you want!) As a baseline of the parameters, let’s use the information from the seal data:

slope = 0.00237, intercept=115.767, sigma = 5.6805, range of seal ages = 958 to 8353, or, if you prefer, seal ages ∼ N(3730.246, 1293.485). Your call what distribution to use for seal age simulation.

Extra credit 1 - test whether the distribution of ages alters power: 3 points

Extra Credit 2 Choose just one of the above elements to vary. Using likelihood to fit models, repeat your power analysis for a chi-square likelihood ratio test. You can use glm(), bbmle or some other means of fitting and obtaining a LRT at your discretion. 5 points.  

```{r}

sim_parameters <- data.frame(intercept = c(100, 115,120,125),
                      slope= 0.00237) %>%
  crossing(residual_sd = 3:8) %>%
  crossing(sample_size = 5:15) %>%    #looking at sample size, intercept, residual
  #set up sampling 
  group_by(intercept, slope, residual_sd, sample_size) %>%
  expand(reps= 1:sample_size) %>%
  ungroup() %>%
  #replicate for simulations 
  crossing(sim= 1:100) %>% 
  mutate(age.days= runif(n(), 958, 8353)) %>% 
           mutate(length.cm = rnorm(n(), intercept + slope* age.days,
                                    residual_sd)) %>%
  ungroup()
  #fit models   
  

 #lim_fit <- function(sim_pop) { 
  lm_fit <- sim_parameters %>% 
    group_by(sim, intercept, slope, residual_sd, sample_size) %>%
    nest() %>%
    mutate(mod = purrr:::map(data, ~lm(length.cm ~ age.days, data=.))) %>%
    mutate(coefs = purrr::map(mod, ~broom::tidy(.))) %>%
    unnest(coefs) %>%
    ungroup() %>%
    filter(term == "age.days")
  
  #pow <- lm_fit %>% 
   # crossing(alpha= c(0.10, 0.05, 0.01)) %>% 
    #group_by(intercept, slope, residual_sd, sample_size, alpha)%>%
    #summarize(power=1-sum(p.value>alpha)/n())%>%
    #ungroup() %>%
    #mutate(alpha = factor(alpha)) 
  
#pow 
  
#power for intercept change 
 pow_intercept <- lm_fit %>% 
   crossing(alpha= c(0.001, 0.01, 0.05, 0.1))%>%
   group_by(intercept, alpha)%>% 
   mutate(power=1-sum(p.value>alpha)/n()) %>%
   ungroup()

###Rinse and repeat for residual sd, and sample size
pow_resid <- lm_fit %>% 
  crossing(alpha=c(0.001, 0.01, 0.05, 0.1)) %>%
  group_by(residual_sd, alpha) %>%
  mutate(power= 1- sum(p.value>alpha)/n()) %>%
  ungroup()

pow_samp_siz <- lm_fit %>% 
  crossing(alpha= c(0.001, 0.01, 0.05, 0.1))%>%
  group_by(sample_size, alpha) %>% 
  mutate(power=  1- sum(p.value>alpha)/n()) %>%
  ungroup()

#plot 
inter_plot <- ggplot(data = pow_intercept, 
       mapping = aes(x = intercept, y = power, color = factor(alpha))) +
  geom_point() +
  geom_line() +
  theme_bw()
resid_plot<- ggplot(data = pow_resid, 
       mapping = aes(x = residual_sd, y = power, color = factor(alpha))) +
  geom_point() +
  geom_line() +
  theme_bw()
samp_siz_plot <- ggplot(data = pow_samp_siz, 
       mapping = aes(x = sample_size, y = power, color = factor(alpha))) +
  geom_point() +
  geom_line() +
  theme_bw()

#visualize#
samp_siz_plot
resid_plot
inter_plot

```
Change in intercept did not have a significant effect on power, increasing residual variance had a negative effect on power and larger sample sizes increased power. 

## 4. I’ve referenced the following figure a few times. I’d like you to demonstrate your understanding of Bayes Theorem by hand showing what the probability of the sun exploding is given the data. Assume that your prior probability that the sun explodes is p(Sun Explodes) = 0.0001. The rest of the information you need is in the cartoon!

```{r}
P_E <- 0.0001  #data points
P_N <- 1- P_exp #probabilit it doesn't explode
P_F <- 1/36    #probabiliy it lies, also prob the it lies given sun explodes
P_T <- 1-P_F  #probabiliy it tells the truth, prob it tells truth given sun explodes

#Prob of explostion give true= P(T | E)* (P(E))/ P(T)
#Prob of it being true P(T)= P(T|E) P(E) + P(T|N) * P(N)



#P(sun explodes | yes) <- P(yes|explodes)*p(Explodes)/Prob(Yes)
#Prob(Yes)= P(Says Ex | explodes)*P(E) + P(Says ex|doesnt)* P(N)

P_yes <- P_T*P_E+P_F*P_N

P_exp_given_data <- (P_T*P_E)/P_yes
P_exp_given_data
```
The probability the sun explodes given the data is 0.0035.  

##5 Quailing at the Prospect of Linear Models
I’d like us to walk through the three different ‘engines’ that we have learned about to fit linear models. To motivate this, we’ll look at Burness et al.’s 2012 study "Post-hatch heat warms adult beaks: irreversible physiological plasticity in Japanese quail 

```{r}
library(readr)

morph <- read.csv("./Data/Morphology data.csv")
morph <- morph[-c(2:5, 8:10)] %>% na.omit()
```

###5.1 Three fits (10 points)
To begin with, I’d like you to fit the relationship that describes how Tarsus (leg) length predicts upper beak (Culmen) length. Fit this relationship using least squares, likelihood, and Bayesian techniques. For each fit, demonstrate that the necessary assumptions have been met. Note, functions used to fit with likelihood and Bayes may or may not behave well when fed NAs. So look out for those errors.
```{r}
#fit with least squares
morph_mod <- lm(morph$Culmen..mm.~ morph$Tarsus..mm.)
#check assumptions 
plot(morph_mod, which=1)   
plot(morph_mod, which=2) #qq-plot looks a little weird on ends 
#tests
anova(morph_mod)
summary(morph_mod)
#log transform 
morph_mod_log <- lm(log(Culmen..mm.)~ Tarsus..mm., data= morph)
#check assumptions again 
plot(morph_mod_log, which=1)
plot(morph_mod_log, which=2) #looks a little better 
#tests
summary(morph_mod_log)  #t-test
anova(morph_mod_log)  #F test
```

```{r}
#fit with likelihood
lik_fun <- function(slope, intercept, resid_sd){
  #data generating process
 fit <- intercept + slope * morph$`Tarsus (mm)`
  
  #likelihood
  sum(dnorm(morph$Culmen..mm., fit, resid_sd, log=TRUE))
}

grid_samp <- crossing(intercept = seq(-0.15, 0.05, .01),
                      slope = seq(0.25,0.45,.01),
                      resid_sd = seq(1.0, 3.0, .01)) %>%
  rowwise() %>%
  mutate(logLik = lik_fun(slope, intercept, resid_sd)) %>%
  ungroup()

#the ML estimates are
grid_samp %>% filter(logLik == max(logLik))

library(dplyr)
library(tidyr)
grid_samp <- crossing(intercept = seq(-0.1,0.1, 0.01),
                      slope = seq(0.1, 0.5, 0.01),
                      resid_sd = seq(1.0,2.0, .05)) %>%
  rowwise() %>%
  mutate(logLik = lik_fun(slope, intercept, resid_sd)) %>%
  ungroup()

#the ML estimates are
grid_samp %>% filter(logLik == max(logLik))
#whelp this isn't fucking working 
```

```{r}
#Bayesian 
library(brms)
library(tidybayes)
library(bayesplot)

morph_lm_bay <- brm(Culmen..mm.~Tarsus..mm.,
                      family = gaussian(link = "identity"),
                      data= morph,
                    file= "./morph_lm_bay")
#test assumptions 
morph_posterior <- posterior_samples(morph_lm_bay, 
                                    add_chain = T)
rhat(morph_lm_bay) #yes! close to 1
mcmc_acf(morph_posterior) #also looks good 
morph_fit <- fitted(morph_lm_bay) %>% as.data.frame() 
morph_res <- residuals(morph_lm_bay) %>% as.data.frame()

qqnorm(morph_res$Estimate) #looks good 
```

5.2 Three interpretations (10 points)
OK, now that we have fits, take a look! Do the coefficients and their associated measures of error in their estimation match? How would we interpret the results from these different analyses differently? Or would we? Note, confint works on lm objects as well.
```{r}
library(tidyverse)
#least squares linear regression
summary(morph_mod)%>% broom::tidy()
confint(morph_mod)

#likelihood

#Baysian 
summary(morph_lm_bay, digits = 4) #so far so good
```

We would interpret Bayesian that it is very likely that Tarsus length affects culmen size. 
We would interpret the least squares as it was very plausible that we can reject the null and we find that it is plausible that the variation is due to correlation and not noise.   
###5.3 Everyday I’m Profilin’ (10 points)For your likelihood fit, are your profiles well behaved? For just the slope, use grid sampling to create a profile. You’ll need to write functions for this, and use the results from your glm() fit to provide the reasonable bounds of what you should be profiling over (3SE should do). Is it well behaved? Plot the profile and give the 80% and 95% CI. Verify your results with profileModel.



###5.4 The Power of the Prior (10 points)
####This data set is pretty big. After excluding NAs in the variables we’re interested in, it’s over 766 lines of data! Now, a lot of data can overhwelm a strong prior. But only to a point. Show first that there is enough data here that a prior for the slope with an estimate of 0.4 and a sd of 0.01 is overwhelmed by the data by demonstrating that it produces similar results to our already fit flat prior.
```{r}
prior_bayes <- function(y, lambda_est, 
                       min_prior=2, max_prior= 20) {
  #DataGenProcess
  lambda <- lambda_est
  
  #prior typically a probability distribution, flat usually assigns equal prob 
  #pois has a lower bound- can't be below 0)
  prior <- dunif(lambda_est,min_prior, max_prior)  #dunif takes min and max 
                          #set max to be weakly informed prior based on data
                          #based on what we know about poisson and our data

  #likelihood 
  lik <- dpois(y, lambda)

  #
  return(prod(lik)*prior)
  }
```

####Second, see if a very small sample size (n = 10) would at least include 0.4 in it’s 95% Credible Interval. 

####Last, demonstrate at what sample size that 95% CL first begins to include 0.4 when we have a strong prior. How much data do we really need to overcome our prior belief? Note, it takes a long time to fit these models, so, try a strategy of spacing out the 3-4 sample sizes, and then zoom in on an interesting region.

###6. Extra CreditMake an election forecast as discussed at https://biol607.github.io/extra.html - but this isn’t just a winner prediction. 1 point for the correct winner. 5 points for correctly predicting the popular vote and being within 10% (3% just for trying!). 5 points for predicting the electoral college and geting no more than 5 states wrong (3 points just for trying). 5 points for predicting the senate races getting no more than 5 states wrong (3 points just for trying). 1 extra point for each vote percentage within your 80% Confidence/Credible Interval. Ditto for the house races.



